Lynott et al. (2014) - Potential error in statistics reporting

### 2017-11-09 manuscript.Rmd
Updated manuscript
* Copy-pasted latest version of Google doc (09-11-2017)
* Fixed minor issues and spelling errors

For an alpha level of 5%, we can conclude statistical equivalence whenever a 90% confidence interval around the observed estimate does not contain the smallest effect size of interest (SESOI). A 90% confidence interval is used because one-sided tests are performed, and it is not necessary to control for multiple comparisons when performing two one-sided tests because both one-sided tests need to be statistically significant to conclude equivalence. Note that although we will use an alpha of 5% in all examples because this alpha level was used in the papers we reanalyze, researchers should carefully decide upon a desired Type 1 error rate, and justify their choice (Lakens et al., 2017). Figure 1 illustrates the possible conclusions we can draw from null hypothesis, minimal effect, and equivalence tests. Although thinking about equivalence tests in terms of the values that can be rejected based on a confidence interval might be the easiest conceptualization, for any given study, you can also think of an equivalence test as determining whether effect sizes or test statistics are closer to zero than some critical, or even whether the *p*-value for a NHST is larger than some *p*-value bound.

@@

Equivalence tests are performed against a value that is considered the smallest effect size of interest (SESOI). Although in the examples here, we use symmetrical bounds around 0 (e.g., -0.3, 0.3), based on for example an explicit cost-benefit analysis, one might use asymmetric bounds (e.g., -0.2, 0.3). The SESOI can sometimes be determined objectively, for example based on just noticeable differences . Most often, however, it is a subjective decision that varies across researchers, fields, and time. In lieu of objective justifications, the SESOI should ideally be based on a cost-benefit analysis. Since both costs and benefits are necessarily subjective, the SESOI will depend on the researcher who designs the study. The goal of setting a SESOI is to clearly justify why designing a study that has a high probability of rejecting effects larger than a specified value contributes to our knowledge base. A SESOI should be chosen such that inferences based on it answer a meaningful question.

@@

We distinguish between three categories of subjective justifications for SESOI. First, researchers can use benchmarks. For example, one might set the SESOI to a standardized effect size of d = 0.5, which would allow one to reject effect as large or larger than a 'medium' effect size (Cohen, 1988). Similarly, effect sizes smaller than a Cohen's d of 0.1 are sometimes considered trivially small (Maxwell, Lau, & Howard, 2015). Relying on a benchmark is the weakest possible justification of a SESOI, and should be avoided. Based on a review of 112 meta-analyses, Weber and Popova (2012) conclude that setting a SESOI in communication research to a medium effect size (*r* = 0.3, or d = 0.5) corresponds to rejecting the highest 25% of effect sizes reported in communications research, and Hemphill (2003) suggests that in psychology a SESOI of d = 0.5 would imply rejecting effects as large as the upper 33% of effect sizes reported in the psychological literature.

@@

Third, researchers can set the SESOI based on the sample size they are planning to collect. The amount of data you can collect limits the inferences you can make. Given the alpha level and the sample size, researchers can calculate the smallest effect size that they have sufficient power to detect. This approach can be used when there are no quantitative predictions made by a theory, and no previous studies have been performed, but researchers *can* justify their sample size. In essence, this approach allows researchers to reject the presence of effects that can be reliably detected given a specific sample size. For example, a researcher who plans to perform a two-sided one-sample *t*-test using an alpha of 5%, based on data from 100 observations, has 90% power to detect an effect of d = 0.364. Using equivalence bounds of d = -0.364 and d = 0.364 in an equivalence test allows one to reject effects the study had high statistical power to detect. An equivalence test based on this approach does not answer any theoretical question (after all, the equivalence bounds are not based on any theoretical predictions) but it answers a *resource* question: Can effects be rejected that can be reliably detected given a specific sample size? In this case, concluding equivalence would suggest one can reject effects that can reasonably be detected with a sample size of 100, and that future studies probably need larger sample sizes to examine a specific question. Whether or not the answer such an equivalence test provides is interesting depends on the sample sizes used in the study, and in part on how typical the chosen sample size is for the research area. One should not expect an equivalence test based on 15 observations testing against a SESOI of d = -1 and d = 1 to substantially increase our knowledge, given that most effects in psychology are substantially smaller than these bounds and require larger sample sizes to be detected. By transparently reporting the effects one can detect and reject, based on the study design, researchers can communicate the information their study contributes, and provide a starting point for a discussion about what a reasonable SESOI is. 
```{r}
p <- 0.05
N <- 100
crit_d <- abs(qt(p/2, (N*2)-2))/sqrt(N/2)
```
#Raw versus standardized equivalence bounds

@@

@@

Banerjee, Chatterjee, & Sinha (2012) reported that `r orig.N` participants who had been asked to describe an unethical deed from their past judged the room to be darker than participants who had been asked to describe an ethical deed ($M_{unethical}= `r orig.m1`$, ${SD}_{unethical}= `r orig.sd1`$, $M_{ethical}=`r orig.m2`$, ${SD}_{ethical}=`r orig.sd2`$, $t(`r orig.df`)= `r orig.t`$, $p= `r orig.p`$, $d= `r orig.d`$). A close replication by Brandt, IJzerman, & Blanken (2014) with `r rep.n1+rep.n2` participants found no significant effect ($t(`r rep.df`)=`r rep.t`$, $p=`r rep.p`$, $d=`r rep.d`$). Following the small telescopes approach, we can calculate the effect size the original study had $33\%$ power to detect --- $d = `r round(d.33, 2)`$ --- and use this as our SESOI. If we run a TOST with Welch's *t*-test for independent samples and equivalence bounds of ${\Delta}_L=`r round(-d.33, 2)`$ and ${\Delta}_U=`r round(d.33, 2)`$, we indeed find that the effect reported by the replication study is statistically equivalent, $t(`r round(Brandt$TOST_df, 2)`)=`r round(min(Brandt$TOST_t1, Brandt$TOST_t2), 2)`$, $p=`r round(max(Brandt$TOST_p1, Brandt$TOST_p2), 3)`$.

@@

```{r, fig.width=6}
source("SEQ_Hyde_Example.R")
```
Hyde, Lindberg, Linn, Ellis, and Williams (2008) report that effect sizes for gender differences in mathematics tests across the 7 million students in the US represent trivial differences, where the authors specify a trivial difference as an effect size smaller than d = 0.1. For example, in grade 3 the difference is d = 0.04, with a standard error of 0.002. When we perform equivalence tests on the meta-analytic effect sizes of IQ difference for grades 2 to 11 (using an alpha level of 0.005 to correct for multiple comparisons), using equivalence bounds of d = -0.1 and d = 0.1, we see that indeed, all effect size estimates are measured with such high precision, we can conclude that they fall within the equivalence bound of d = -0.1 and d = 0.1, and can be considered trivially small, according to the bounds set by the authors (figure 2B). However, note that all of the effects are also statistically different from zero, as one might expect when there is no random assignment to conditions, and samples sizes are huge. This is one way in which the use of equivalence tests improve hypothesis testing procedures in science, by distinguishing between statistical significance and practical significance.

@@

We find that the TOST is non-significant, t(`r round(Moon$TOST_df, 2)`) = `r round(min(abs(c(Moon$TOST_t1, Moon$TOST_t2))), 2)`, p = `r round(max(Moon$TOST_p1, Moon$TOST_p2), 2)` (figure 2C). Thus, we cannot reject the possibility that the true effect really is larger than our SESOI.[^footnote] When results are neither statistically different from zero nor statistically equivalent, there is insufficient data to draw conclusions. Further studies are needed, which can be analyzed using a (small-scale) meta-analysis. The additional data will narrow the confidence interval around the observed effect, allowing us to reject the null, reject the SESOI, or both. Analogous to the large sample sizes needed to detect small effects, a lot of data is needed to reject the SESOI when very narrow bounds are used for the equivalence test (e.g., $\Delta_{lower}$ = -0.1 and $\Delta_{upper}$ = 0.1).

@@

Lynott et al. (2014) conducted a study to investigate the effect of being exposed to physically warm or cold stimuli on subsequent judgments related to interpersonal warmth and prosocial behavior (replicating Williams and Bargh, 2008). They observed that `r selfish.cold * 100`\% of participants who received a cold pack ($n_{cold}=`r n.cold`$) opted to receive a reward for themselves, while `r selfish.hot * 100`\% of participants who received a warm pack ($n_{hot}=`r n.hot`$) did the same. In a z-test for the difference between the proportions, this effect is not statistically significant ($Diff=`r Lynott$dif * 100`$\%, $Z=`r Lynott$NHST_z`$, $p=`r Lynott$NHST_p`$). Having obtained a non-significant result, we are left with the same two  alternatives as in example 3.

In this case, it is not really clear what an objective criterion for the smallest effect size of interest could be. However, since this is a replication, we could decide that we do not care about an effect if the difference is smaller than the smallest effect that the original study could detect. This essentially means that we use the critical z value (~1.96 in a two-tailed test with an alpha of 0.05) as our bounds ($\Delta = +/- round(qnorm(1-0.05/2),2)$). To figure out what difference corresponds to a critical z in the original study, we multiply critical z with the standard error:

Having determined the SESOI, we may also decide that since this is a replication, we are only interested in whether the effect of receiving a cold pack is significantly *lower* than the smallest detectable effect of the original study. In this case, we can still use a TOST to analyse the data, but we only need to consider the test against $\Delta_{upper}$ because an effect with CIs crossing $\Delta_{lower}$ can still be *lower* than the original effect.

We find that the test against $\Delta_{upper}$ is `r ifelse(Lynott$TOST_p2 < Lynott$alpha, "significant", "non-significant")`, $z =$ `r round(Lynott$TOST_z2, 2)`, $p =$ `r ifelse(round(Lynott$TOST_p2, 3) >= 0.001, paste("=", round(Lynott$TOST_p2, 2)), "< 0.001")` (see figure 2D). Thus, we can conclude that the statistically non-significant effect is also statistically inferior to our SESOI.

@@

```{r include=FALSE}
source("examples_plotgrid.R")
```
```{r echo=FALSE}
plot(example.grid)
```

#Adressing Trello cards

For the example where the results are inconclusive, explain how adding more data, will, assuming the true effect is 0, lead to a smaller confidence interval, and thus will lead to an informative test. However, with increasing data, the effect could also slightly increase, to above or at the SESOI level, and then the effect will be significant and not equivalent. With massive amounts of data, many results will be significant and equivalent.

@@

The given examples and figures XXX to YYY illustrate that equivalence testing is a relatively simple and flexible technique: As long as we can calculate a confidence interval around a mean effect, we can compare it to one or several bounds. In that sense, the result of an equivalence test can be obtained by mere visual inspection of the CI (Seaman & Serlin, 1998; Tryon, 2001). This ease of use might tempt researchers to set equivalence bounds *after* looking at their data.
For instance, in example 3 we showed that Moon & Roeder's replication of the identity-priming study by Shih et al. (1999; example 3) does not exclude effects that are large enough to affect a student's final grade on a test. However, if we look at the $90 \%$ CI (figure XXX), we can see that the data would have been statistically equivalent if the bounds had been set to  a slightly larger SESOI, e.g. a mean difference of 0.1.
It is important to point out that this strategy --- setting equivalence bounds *based on the obtained results* --- may be considered as "boundary hacking": Regardless of the underlying true effect, any finite CI will exclude some values, which means that this procedure will lead to capitalization on chance.

In order to avoid fooling ourselves more often than $\alpha \%$ of the time, equivalence bounds should be specified before results are known (ideally as part of a preregistration; cf. CONSORT guidelines, Piaggio et al., 2006). A perhaps helpful guiding question may be: "Will testing against these bounds yield interesting information in the long run?"

#Justify your equivalence bounds

Like with any statistical test, the usefulness of the result an equivalence test gives us depends on the question we ask. The question manifests in the bounds we set: Is our data surprising if the true effect lies outside of the bounds? If we set very wide bounds, we ask if our data is surprising given that the true effect is very large[^footnote]. For most psychological effects, the answer will usually be “yes,” which means that it may not be a very interesting question to ask. In order to obtain a meaningful result, we have to set the bounds based on meaningful criteria: Our SESOI should really be *the smallest effect we are interested in* for a given study.  For this reason, researchers should provide a detailed justification of the choice for a SESOI. Some studies in the literature simply state the SESOI that is used, without providing a justification. For example, Brown and colleagues (2017) and Schumann and colleagues (2017)use a SESOI of Cohen’s d = +/- 0.5, but it is not clear why these bounds were chosen[^footnote]. An equivalence test should answer a question, and justifying the SESOI is an essential part of asking the question. An equivalence test of d = 0.5 might very well answer a question the researchers are interested in (for one possible justification based on minimally important differences, see Norman et al., 2003), but the reason this SESOI was chosen should be explicitly justified.

[^footnote]: To be precise, wide bounds set symmetrically around 0 ask if the data is surprising given that the true effect is very *extreme*, i.e. either very large (positive) or very small (negative).
[^footnote]: In these two cases, one could perhaps guess that benchmark values for a Cohen’s d of “medium” size was chosen.

>>>>

For an alpha level of 5%, we can conclude statistical equivalence whenever a 90% confidence interval around the observed estimate does not contain the smallest effect size of interest (SESOI). A 90% confidence interval is used because one-sided tests are performed, and it is not necessary to control for multiple comparisons when performing two one-sided tests because both tests need to be statistically significant to conclude equivalence. Note that although we will use an alpha of 5% in all examples because this alpha level was used in the papers we reanalyze, researchers should carefully decide upon a desired Type 1 error rate, and justify their choice (Lakens et al., 2017). Figure 1 illustrates the possible conclusions we can draw from null hypothesis, minimal effect, and equivalence tests. Although thinking about equivalence tests in terms of the values that can be rejected based on a confidence interval might be the easiest conceptualization, for any given study, you can also think of an equivalence test as determining whether effect sizes or test statistics are closer to zero than some critical, or even whether the *p*-value for a NHST is larger than some *p*-value bound.

@@

Equivalence tests are performed against a value that is considered the smallest effect size of interest (SESOI). Although we use symmetrical bounds around 0 in the examples we will discuss below (e.g., $\Delta_{L} = -0.3$, $\Delta_{U} = 0.3$), it is possible to use asymmetric bounds, for example based on an explicit cost-benefit analysis (e.g., $\Delta_{L} = -0.2$, $\Delta_{U} = 0.3$). The SESOI can sometimes be determined objectively, for example based on just noticeable differences . Most often, however, it is a subjective decision that varies across researchers, fields, and time. In lieu of objective justifications, the SESOI should ideally be based on a cost-benefit analysis. Since both costs and benefits are necessarily subjective, the SESOI will depend on the researcher who designs the study. The goal of setting a SESOI is to clearly justify why designing a study that has a high probability of rejecting effects larger than a specified value contributes to our knowledge base. A SESOI should be chosen such that inferences based on it answer a meaningful question.

@@

We distinguish between three categories of subjective justifications for the SESOI. First, researchers can use benchmarks. For example, one might set the SESOI to a standardized effect size of d = 0.5, which would allow one to reject effect as large or larger than a 'medium' effect size (Cohen, 1988). Similarly, effect sizes smaller than a Cohen's d of 0.1 are sometimes considered trivially small (Maxwell, Lau, & Howard, 2015). Relying on a benchmark is the weakest possible justification of a SESOI, and should be avoided. Based on a review of 112 meta-analyses, Weber and Popova (2012) conclude that setting a SESOI in communication research to a medium effect size (*r* = 0.3, or d = 0.5) corresponds to rejecting the highest 25% of effect sizes reported in communications research, and Hemphill (2003) suggests that in psychology a SESOI of d = 0.5 would imply rejecting effects as large as the upper 33% of effect sizes reported in the psychological literature.

@@

Third, researchers can set the SESOI based on the sample size they are planning to collect. The amount of data you can collect limits the inferences you can make. Given the alpha level and the sample size, researchers can calculate the smallest effect size that they have sufficient power to detect[^footnote]. This approach can be used when there are no quantitative predictions made by a theory, and no previous studies have been performed, but researchers *can* justify their sample size. In essence, this approach allows researchers to reject the presence of effects that can be reliably detected given a specific sample size. For example, a researcher who plans to perform a (two-sided) one-sample *t*-test using an alpha of 5%, based on data from 100 observations, has 90% power to detect an effect of d = 0.364. Using equivalence bounds of $\Delta_{L} = -0.364$ and $\Delta_{U} = 0.364$ in an equivalence test allows one to reject effects the study had high statistical power to detect. An equivalence test based on this approach does not answer any theoretical question (after all, the equivalence bounds are not based on any theoretical predictions) but it answers a *resource* question: Can effects be rejected that could reliably be detected given a specific sample size? In this case, concluding equivalence would suggest that effects which could reasonably be detected with a sample size of 100 can be rejected, and that future studies probably need larger sample sizes to examine a specific question. Whether or not the answer such an equivalence test provides is interesting depends on the sample sizes used in the study, and in part on how typical the chosen sample size is for the research area. An equivalence test based on 15 observations testing against a SESOI of d = 1 cannot be expected to substantially increase our knowledge, given that most effects in psychology are substantially smaller than these bounds and require larger sample sizes to be detected. By transparently reporting the effects one can detect and reject, based on the study design, researchers can communicate the information their study contributes, and provide a starting point for a discussion about what a reasonable SESOI is. 
```{r}
p <- 0.05
N <- 100
crit_d <- abs(qt(p/2, (N*2)-2))/sqrt(N/2)
```

[^footnote]: This approach is conceptually very similar to the *power approach*, where the effect size you had 95% power to detect is calculated, and the presence of effects more extreme than this value is rejected after observing a *p*-value *larger than* 0.05 in a traditional NHST. However, Meyners (2012) explains this approach is not recommended (even though it is common) because it ignores the possibility that effects are significant *and* equivalent, and the approach does not accurately control error rates.

#Raw versus standardized equivalence bounds

@@

In the remainder of this article we will provide five detailed examples of equivalence tests performed on published studies. These concrete and easy to follow examples will illustrate all approaches to setting equivalence bounds discussed above, and demonstrate how to perform and report equivalence tests. The code to reproduce these analyses is available at XXXX.XXXX.

```{r include=FALSE}
source("examples_plotgrid.R")
```
```{r echo=FALSE}
#  Plot the TOST examples
plot(example.grid)
```

Figure caption: Figure 2. Effects with TOST CIs (thick lines) and NHST CIs (thin lines), NHST H0 (solid vertical line) and TOST H0 (dashed vertical lines) displayed. **A)** Example 1 - The mean difference. **B)** Example 2 - The effects included in the meta-analysis. **C)** Example 3 - The mean difference. **D)** - The proportion difference.

@@

Banerjee, Chatterjee, & Sinha (2012) reported that `r orig.N` participants who had been asked to describe an unethical deed from their past judged the room to be darker than participants who had been asked to describe an ethical deed ($M_{unethical}= `r orig.m1`$, ${SD}_{unethical}= `r orig.sd1`$, $M_{ethical}=`r orig.m2`$, ${SD}_{ethical}=`r orig.sd2`$, $t(`r orig.df`)= `r orig.t`$, $p= `r orig.p`$, $d= `r orig.d`$). A close replication by Brandt, IJzerman, & Blanken (2014) with `r rep.n1+rep.n2` participants found no significant effect ($t(`r rep.df`)=`r rep.t`$, $p=`r rep.p`$, $d=`r rep.d`$). Following the small telescopes approach, we can calculate the effect size the original study had $33\%$ power to detect --- $d = `r round(d.33, 2)`$ --- and use this as our SESOI. If we run a TOST with Welch's *t*-test for independent samples and equivalence bounds of ${\Delta}_L=`r round(-d.33, 2)`$ and ${\Delta}_U=`r round(d.33, 2)`$, we indeed find that the effect reported by the replication study is statistically equivalent, $t(`r round(Brandt$TOST_df, 2)`)=`r round(min(Brandt$TOST_t1, Brandt$TOST_t2), 2)`$, $p=`r round(max(Brandt$TOST_p1, Brandt$TOST_p2), 3)`$ (Figure 2A).

@@

```{r, include=FALSE}
source("SEQ_Hyde_Example.R")
```

Hyde, Lindberg, Linn, Ellis, and Williams (2008) report that effect sizes for gender differences in mathematics tests across the 7 million students in the US represent trivial differences, where the authors specify a trivial difference as an effect size smaller than d = 0.1. For example, in grade 3 the difference is d = 0.04, with a standard error of 0.002. When we perform equivalence tests on the meta-analytic effect sizes of IQ difference for grades 2 to 11 (using an alpha level of 0.005 to correct for multiple comparisons), using equivalence bounds of d = -0.1 and d = 0.1, we see that indeed, all effect size estimates are measured with such high precision, we can conclude that they fall within the equivalence bound of d = -0.1 and d = 0.1, and can be considered trivially small, according to the bounds set by the authors (figure 2B). However, note that all of the effects are also statistically different from zero, as one might expect when there is no random assignment to conditions, and samples sizes are huge. This is one way in which the use of equivalence tests improve hypothesis testing procedures in science, by distinguishing between statistical significance and practical significance.

@@

We find that the TOST is non-significant, t(`r round(Moon$TOST_df, 2)`) = `r round(min(abs(c(Moon$TOST_t1, Moon$TOST_t2))), 2)`, p = `r round(max(Moon$TOST_p1, Moon$TOST_p2), 2)` (figure 2C). Thus, we cannot reject the possibility that the true effect really is larger than our SESOI.[^footnote] When results are neither statistically different from zero nor statistically equivalent, there is insufficient data to draw conclusions. Further studies are needed, which can be analyzed using a (small-scale) meta-analysis. The additional data will narrow the confidence interval around the observed effect, allowing us to reject the null, reject the SESOI, or both. Analogous to the large sample sizes needed to detect small effects, a lot of data is needed to reject the SESOI when very narrow bounds are used for the equivalence test (e.g., $\Delta_{L}$ = -0.1 and $\Delta_{U}$ = 0.1).

@@

Lynott et al. (2014) conducted a study to investigate the effect of being exposed to physically warm or cold stimuli on subsequent judgments related to interpersonal warmth and prosocial behavior (replicating Williams and Bargh, 2008). They observed that `r round(selfish.cold,2)*100`\% of participants who received a cold pack ($n_{cold}=`r round(n.cold,2)`$) opted to receive a reward for themselves, while `r round(selfish.hot,2)*100`\% of participants who received a warm pack ($n_{hot}=`r n.hot`$) did the same. In a z-test for the difference between the proportions, this effect is not statistically significant ($Diff=`r round(Lynott$dif * 100)`$\%, $Z=`r round(Lynott$NHST_z,2)`$, $p=`r round(Lynott$NHST_p,2)`$). Having obtained a non-significant result, we are left with the same two  alternatives as in example 3.

In this case, it is not really clear what an objective criterion for the smallest effect size of interest could be. However, since this is a replication, we could decide that we do not care about an effect if the difference is smaller than the smallest effect that the original study could detect. This essentially means that we use the critical z value (~1.96 in a two-tailed test with an alpha of 0.05) as our bounds ($\Delta = +/- `r round(qnorm(1-0.05/2),2)`$). To figure out what difference corresponds to a critical z in the original study, we multiply critical z with the standard error:

Having determined the SESOI, we may also decide that since this is a replication, we are only interested in whether the effect of receiving a cold pack is significantly *lower* than the smallest detectable effect of the original study. In this case, we can still use a TOST to analyse the data, but we only need to consider the test against $\Delta_{U}$ because an effect with CIs crossing $\Delta_{L}$ can still be *lower* than the original effect.

We find that the test against $\Delta_{U}$ is `r ifelse(Lynott$TOST_p2 < Lynott$alpha, "significant", "non-significant")`, $z =$ `r round(Lynott$TOST_z2, 2)`, $p =$ `r ifelse(round(Lynott$TOST_p2, 3) >= 0.001, paste("=", round(Lynott$TOST_p2, 2)), "< 0.001")` (see figure 2D). Thus, we can conclude that the statistically non-significant effect is also statistically inferior to our SESOI.

@@

## Box 1: Calculating an equivalence test in R

If you want to calculate an equivalence test based on summary statistics, you can use the "TOSTER" package in the open-source programming language R. Using TOSTER requires very little experience with R, or programming in general. Below are the few simple steps required to calculate an equivalence test with TOSTER.

1. Download, install, and open R.
2. Run `install.packages("TOSTER")` in the console.
3. Run `library(TOSTER)` in the console.
4. The example below reproduces the result of example 1 in R. The parameters of the test are defined inside the parentheses, and you can run `?TOSTtwo` in the console to get information about them. Simply copy the example to the console, replace the values with the corresponding values of your own study, and run the code. Results and a plot will automatically be printed.
```{r eval=FALSE, echo=TRUE, tidy=TRUE, tidy.opts=list(width.cutoff=60)} 
TOSTtwo(m1  = 4.7857, m2  = 4.6569,sd1 = 1.0897, sd2 = 1.1895, n1 = 49, n2 = 51, low_eqbound_d = -0.6401696, high_eqbound_d = 0.6401696, alpha = 0.05, var.equal = FALSE)
```

5. The example above uses the `TOSTtwo()` function in the TOSTER package which calculates an equivalence test for an independent samples t-test. If you need to calculate TOST for a different type of test (e.g. a correlation), you can find documentation for `TOSTr()` and other functions of the TOSTER package by going to https://cran.r-project.org/web/packages/TOSTER/TOSTER.pdf in your web browser.

@@

Equivalence testing is a relatively simple and flexible technique to falsify hypotheses. As long as we can calculate a confidence interval around a parameter estimate of interest, we can compare it to a smallest effect size of interest. The result of an equivalence test can be obtained by mere visual inspection of the CI (Seaman & Serlin, 1998; Tryon, 2001), or by performing two one-sided tests.

Equivalence bounds should be specified before results are known, ideally as part of a pre-registration (cf. CONSORT guidelines, Piaggio et al., 2006). Researchers might be tempted to set equivalence bounds *after* looking at their data, but this "boundary hacking" will increase error rates.

As with any statistical test, the usefulness of the result of an equivalence test depends on the question we ask. The question manifests itself in the bounds we set: Is our data surprising if the true effect lies outside of the bounds? If we test against very wide bounds, observing statistical significance can hardly be considered surprising, given that most effects in psychology are small to medium (Hemphill, 2003). Examining the papers citing Lakens (2017), we see some researchers state a SESOI without providing a justification for it (Brown and colleagues, 2017; Schumann and colleagues, 2017)[^footnote]. An equivalence test using a SESOI of d = 0.5 might very well answer a question the researchers are interested in (for one possible justification based on minimally important differences, see Norman et al., 2003), but researchers should always provide a detailed justification of their chosen SESOI, given that this is in essence a justification for why they ask a specific question from their data.

For many researchers, the biggest challenge in designing studies where they will perform both NHST and equivalence tests by default will be the need to specify the smallest effect size of interest. Psychological theories are usually too vague to derive precise predictions, and if there are no theoretical reference points, natural constraints, or prior studies a researcher can use to define the SESOI for a new hypothesis, any choice will be arbitrary to some extent. In some research lines researchers might simply use equivalence tests to reject consecutively smaller effect sizes by performing studies with increasingly larger sample sizes while controlling error rates, until no one is willing to invest the time and resources needed to examine the presence of even smaller effects. Nevertheless, it is important to realize that not specifying a SESOI in our research questions will severely hinder theoretical progress. While our initial choices may be arbitrary and subjective, our goal should be to use them as starting points to develop better and more falsifiable theories over time, and ultimately be able to make more precise predictions than we can today.

[^footnote]: In these two cases, one could perhaps guess that benchmark values for a Cohen’s d of “medium” size was chosen.
